---
title: Estimate ctmm movement model and create seasonal range KDEs
---

## Introduction

Reference: https://ctmm-initiative.github.io/ctmmwebdoc/articles/package_usage.html

```{r}
#| message: FALSE
#| warning: FALSE
library(sf)
library(terra)
library(dplyr)
library(ctmm)
library(magrittr)
library(ctmmweb)
library(data.table)

#load data and convert to telemetry object for ctmm functions
ccgps_summer_filtered <- read.csv('../data/yt/gps_fallrut.csv')

fall_telem <- as.telemetry(ccgps_fall_filtered)
```

## Variograms

Fall Rut. Caribou 38857 has a gap, 38873 monitored for 8 days. Others have ~15 - 20 days (short season)

```{r}
#| message: FALSE
#| warning: FALSE
#| fig-width: 9
#| fig-height: 9
f_vario_list <- lapply(fall_telem, ctmm::variogram)
# names of vario_list are needed for figure titles
names(f_vario_list) <- names(fall_telem)
## 1 - 6
plot_vario(f_vario_list[1:6]) # plot 6 caribou at a time. 27 caribou total. 

## 7 - 12
plot_vario(f_vario_list[7:12])

## 13 - 18
plot_vario(f_vario_list[13:18])

## 19 - 24
plot_vario(f_vario_list[19:24])

## 25 - 27
plot_vario(f_vario_list[25])

```

Compare initial movement model parameter guesses to data via variograms

```{r}
#| message: FALSE
#| warning: FALSE
#| fig-width: 9
#| fig-height: 9
f_guess_list <- lapply(fall_telem,
                     function(tele)
                       ctmm::ctmm.guess(tele, interactive = FALSE))

## 1- 6
plot_vario(f_vario_list[1:6], f_guess_list[1:6])

## 7 - 12
plot_vario(f_vario_list[7:12], f_guess_list[7:12])

## 13 - 18
plot_vario(f_vario_list[13:18], f_guess_list[13:18])

## 19 - 24
plot_vario(f_vario_list[19:24], f_guess_list[19:24])

## 25 - 27
plot_vario(f_vario_list[25], f_guess_list[25])
```

Try different models of multiple animals in parallel with ctm.select. Parallel mode can be turned off with parallel = LASE.

`f_model_try_res` holds a list of items named by animal names. Each item hold the attempted models for that animal as a sub list, named by model type.

```{r}
#| message: FALSE
#| warning: FALSE
f_model_try_res <- par_try_models(fall_telem)
f_model_summary <- summary_tried_models(f_model_try_res)
knitr::kable(f_model_summary)
```

Keep only models with lowest dAICc

```{r}
#| message: FALSE
#| warning: FALSE
f_lowest_aic_models <- f_model_summary %>% filter(dAICc==0.0)
knitr::kable(f_lowest_aic_models)
```

Make selecting a subset easier by converting nested list to a flat list.

```{r}
#| message: FALSE
#| warning: FALSE
#| fig-width: 9
#| fig-height: 9
f_model_list <- flatten_models(f_model_try_res)

#we need the model_name to filter the model list, and the animal name to filter the variograms
f_names_sub2 <- f_model_summary[(model_no %in% f_lowest_aic_models$model_no), .(model_name, identity)]

# filter model list by model names to get subset of model list.
f_model_list_sub2 <- f_model_list[f_names_sub2$model_name]

# get corresponding variograms by animal names.
f_vario_list_sub2 <- f_vario_list[f_names_sub2$identity]
# specify a different color for model
## 1 - 6
plot_vario(f_vario_list_sub2[1:6], f_model_list_sub2[1:6], model_color = "purple")

## 7 - 12
plot_vario(f_vario_list_sub2[7:12], f_model_list_sub2[7:12], model_color = "purple")

## 13 - 18
plot_vario(f_vario_list_sub2[13:18], f_model_list_sub2[13:18], model_color = "purple")

## 19 - 24
plot_vario(f_vario_list_sub2[19:24], f_model_list_sub2[19:24], model_color = "purple")

## 25 - 27
plot_vario(f_vario_list_sub2[25], f_model_list_sub2[25], model_color = "purple")

```

## Home Range

## Batch calculate home ranges 

Use when lots of individuals and don't need to fine tune each model (this case?)
Compared with first calculation of home range and the outputs are the same. 

```{r}
#| message: FALSE
#| warning: FALSE
#| fig-width: 9
#| fig-height: 9

# one line to try models. The result is a nested list of CTMM models
f_model_try_res <- par_try_models(fall_telem)
#save.image("ew_test.RData")
# summary models to find the best
f_model_summary <- summary_tried_models(f_model_try_res)
# select the best model name for each individual, which is the first one with 0 AICc
f_best_model_names <- f_model_summary[, .(best_model_name = model_name[1]),
                                  by = "identity"]$best_model_name
# get a flat list of model object
f_model_list <- flatten_models(f_model_try_res)
# get the best model objects
f_best_models <- f_model_list[f_best_model_names]
# calculate home range in same grid, with optimal weighting on
f_hrange_list_same_grid <- akde(fall_telem, f_best_models, 
                              weights = TRUE)
# calculate home range separately, with optimal weighting on
f_hrange_list <- par_hrange_each(fall_telem, f_best_models,
                               rep.int(list(TRUE), length(fall_telem)))
## 1 - 6
plot_ud(f_hrange_list[1:6])

## 7 - 12
plot_ud(f_hrange_list[7:12])

## 13 - 18
plot_ud(f_hrange_list[13:18])

## 19 - 24
plot_ud(f_hrange_list[19:24])

## 25 - 27
plot_ud(f_hrange_list[25])
```

## Map all home ranges

```{r}
# A map with home range estimates need: list of home range UD object, vector of level.UD in ctmm::plot.telemetry, a vector of color names for each home range estimate.
range_map(f_hrange_list, 0.95, rainbow(length(f_hrange_list)))

```

## Save home ranges / KDEs

Saving home ranges to a geopackage. 

```{r}
for (i in 1:length(f_hrange_list)) {
  x <- f_hrange_list[i]
  z <- as.sf(x[[1]]) %>% st_transform(3578)
  st_write(z, "../data/yt/fr_hranges.gpkg", paste0("caribou",i), delete_layer=TRUE)
}

```

