--- 
title: "Connectivity Explorer"
subtitle: ""
format: dashboard
---

```{r}
#| context: setup
library(terra)
library(maptiles)
v <- vect('../data/mcp15k.gpkg', 'mcp_buff15k')
bg <- get_tiles(v,
  crop = TRUE, provider = "Esri.WorldShadedRelief",
  zoom = 8, project = TRUE, cachedir = "."
)
```

# {.sidebar}

**Movement**

Home ranges were estimated using the autocorrelated kernel density estimator (aKDE).

**Habitat**

We used three different types of models:

- GLM = Generalized linear model
- GLMM = Generalized linear mixed model
- RFM = Random forest model

<hr>

**Connectivity**

Landscape resistance is based on a negative exponential transformation of the random forest predictive surface, where c is a measure of increasing mobility.

<hr>

**Report**

- [Analysis workbook](../index.html)

# Movement

# Habitat

## Row {height=50%}

### Column {.tabset}

```{r}
#| title: Early winter (GLM)
r <- rast("../data/predict30/glm/earlywintermap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (GLMM)
r <- rast("../data/predict30/glmm/earlywintermap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (RFM)
r <- rast("../data/predict30/randomforest/earlywintermap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

### Column {.tabset}

```{r}
#| title: Late winter (GLM)
r <- rast("../data/predict30/glm/latewintermap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (GLMM)
r <- rast("../data/predict30/glmm/latewintermap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (RFM)
r <- rast("../data/predict30/randomforest/latewintermap.tif")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

## Row {height=50%}

### Column {.tabset}

```{r}
#| title: Summer (GLM)
r <- rast("../data/predict30/glm/summermap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (GLMM)
r <- rast("../data/predict30/glmm/summermap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (RFM)
r <- rast("../data/predict30/randomforest/summermap.tif")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

### Column {.tabset}

```{r}
#| title: Fall rut (GLM)
r <- rast("../data/predict30/glm/fallrutmap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (GLMM)
r <- rast("../data/predict30/glmm/fallrutmap.tif") #|>
  #project("+proj=longlat +datum=WGS84")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

```{r}
#| title: (RFM)
r <- rast("../data/predict30/randomforest/fallrutmap.tif")
plotRGB(bg)
plot(r, breaks=5, breakby='cases', alpha=0.75, add=TRUE)
lines(v, col="black", lwd=2)
```

# Connectivity

## Row {height=50%}

### Column {.tabset}

```{r}
#| title: Early winter (c=0.25)
  r1 = rast("../data/omniscape300/earlywinter_RF300_C25_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=2)
  r1 = rast("../data/omniscape300/earlywinter_RF300_C200_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=8)
  r1 = rast("../data/omniscape300/earlywinter_RF300_C800_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

### Column {.tabset}

```{r}
#| title: Late winter (c=0.25)
  r2 = rast("../data/omniscape300/latewinter_RF300_C25_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r2) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r2, main="Late winter",  alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=2)
  r1 = rast("../data/omniscape300/latewinter_RF300_C200_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=8)
  r1 = rast("../data/omniscape300/latewinter_RF300_C800_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

## Row {height=50%}

### Column {.tabset}

```{r}
#| title: Summer (0.25)
  r3 = rast("../data/omniscape300/summer_RF300_C25_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r3) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r3, main="Summer", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=2)
  r1 = rast("../data/omniscape300/summer_RF300_C200_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=8)
  r1 = rast("../data/omniscape300/summer_RF300_C800_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

### Column {.tabset}

```{r}
#| title: Fall rut (c=0.25)
  r4 = rast("../data/omniscape300/fallrut_RF300_C25_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r4) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r4, main="Fall rut", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=2)
  r1 = rast("../data/omniscape300/fallrut_RF300_C200_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```

```{r}
#| title: (c=8)
  r1 = rast("../data/omniscape300/fallrut_RF300_C800_R1000_T50/categorized_connectivity.tif")
  plotRGB(bg)
  levels(r1) <- data.frame(value=1:4, connectivity=c("Impeded", "Diffuse", "Intensified", "Channelized"))
  plot(r1, main="Early winter", alpha=0.5, add=TRUE)
  lines(v, col="black", lwd=2)
```
