---
title: Home ranges
execute:
  freeze: auto
  echo: true
---

## Introduction

Objectives:

- Autocorrelated home range estimation
- https://ecoisilva.github.io/AKDE_minireview/code/AKDE_R-tutorial.html

```{r}
#| warning: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(ctmm)
```

## Step 1 - Formatting and loading an animal tracking dataset

```{r}
#| warning: false
c1 <- read.csv("../data/yt/gps.csv")
head(c1)
glimpse(c1)
```

**Convert to telemetry object:**

```{r}
#| message: false
#| warning: false
c1t <- as.telemetry(c1)[['43140']]
plot(c1t)
```

## Step 2 - Checking for the range residency assumption

```{r}
level <- 0.95 # we want to display 95% confidence intervals
xlim <- c(0,2 %#% "month") # to create a window of 2 months
SVF <- variogram(c1t)
par(mfrow = c(1,2))
plot(SVF, fraction = 1, level = level)
abline(v = 1, col = "red", lty = 2) # adding a line at 1 month 
plot(SVF, xlim = xlim, level = level)
abline(v = 1, col = "red", lty = 2)
```

## Step 3 - Selecting the best-fit movement model through model selection

**Calculate an automated model guesstimate**

```{r}
GUESS1 <- ctmm.guess(c1t, interactive = FALSE)
# Automated model selection, starting from GUESS
FIT1_ML <- ctmm.select(c1t, GUESS1, method = 'ML')
FIT1_pHREML <- ctmm.select(c1t, GUESS1, method = 'pHREML')
# Reminder: it will default to pHREML if no method is specified.
summary(FIT1_ML)
summary(FIT1_pHREML)
```

**Summarize model**

```{r}
FIT1_pHREML_verbose <- ctmm.select(c1t, GUESS1, verbose = TRUE)
summary(FIT1_pHREML_verbose)
```

## Step 4 - Feeding a movement model into the home range estimator

**Run an area-corrected AKDE:**

```{r}
UD1_ML <- akde(c1t, FIT1_ML)
UD1_pHREML <- akde(c1t, FIT1_pHREML)
summary(UD1_pHREML)$CI # home range area estimation
```

## Step 5 - Evaluating additional biases, applying mitigation measures

```{r}
summary(UD1_pHREML)$DOF["area"] # effective sample size of animal1
nrow(c1t) # absolute sample size
UD1w_pHREML <- akde(c1t, FIT1_pHREML, weights = TRUE)
summary(UD1w_pHREML)$CI # home range area estimation (weighted)
```

**Creating an extent that includes both UDs at the 95% CI level:**

```{r}
EXT <- extent(list(UD1_ML, UD1_pHREML, UD1w_pHREML), level = 0.95)
```

**Plotting pHREML (with and without weights) side-by-side:**

```{r}
par(mfrow = c(1,2))
plot(c1t, UD = UD1_pHREML, ext = EXT)
title(expression("pHREML AKDE"["C"]))
plot(c1t, UD = UD1w_pHREML, ext = EXT)
title(expression("pHREML wAKDE"["C"]))

( 1 - summary(UD1_ML)$CI[1,2] / summary(UD1w_pHREML)$CI[1,2] ) * 100
```
