---
title: "6. Omniscape"
format: gfm
---

```{r}
#| message: FALSE
#| warning: FALSE
library(sf)
library(terra)
library(tidyverse)
library(rpart)
library(rpart.plot)
library(predicts)
library(randomForest)
set.seed(123)
```

# Cumulative current flow

The cumulative current flow map measures the total current flowing through the landscape and can be interpreted as representing forest connectivity. Higher values indicate higher connectivity. In the map below, proposed IPCAs are indicated by transparent blue polygons.

## Early winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m1s1c1 = rast("../../data/omniscape300/earlywinter_RF300_C25_R1000_T50/cum_currmap.tif")
plot(m1s1c1)
```

## c = 2

```{r}
#| echo: false
m1s1c2 = rast("../../data/omniscape300/earlywinter_RF300_C200_R1000_T50/cum_currmap.tif")
plot(m1s1c2)
```

## c = 8

```{r}
#| echo: false
m1s1c3 = rast("../../data/omniscape300/earlywinter_RF300_C800_R1000_T50/cum_currmap.tif")
plot(m1s1c3)
```

:::

## Late winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m1s2c1 = rast("../../data/omniscape300/latewinter_RF300_C25_R1000_T50/cum_currmap.tif")
plot(m1s2c1)
```

## c = 2

```{r}
#| echo: false
m1s2c2 = rast("../../data/omniscape300/latewinter_RF300_C200_R1000_T50/cum_currmap.tif")
plot(m1s2c2)
```

## c = 8

```{r}
#| echo: false
m1s2c2 = rast("../../data/omniscape300/latewinter_RF300_C800_R1000_T50/cum_currmap.tif")
plot(m1s2c2)
```

:::

## Summer

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m1s3c1 = rast("../../data/omniscape300/summer_RF300_C25_R1000_T50/cum_currmap.tif")
plot(m1s3c1)
```

## c = 2

```{r}
#| echo: false
m1s3c2 = rast("../../data/omniscape300/summer_RF300_C200_R1000_T50/cum_currmap.tif")
plot(m1s3c2)
```

## c = 8

```{r}
#| echo: false
m1s3c3 = rast("../../data/omniscape300/summer_RF300_C800_R1000_T50/cum_currmap.tif")
plot(m1s3c3)
```

:::


## Fall rut

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m1s4c1 = rast("../../data/omniscape300/fallrut_RF300_C25_R1000_T50/cum_currmap.tif")
plot(m1s4c1)
```

## c = 2

```{r}
#| echo: false
m1s4c2 = rast("../../data/omniscape300/fallrut_RF300_C200_R1000_T50/cum_currmap.tif")
plot(m1s4c2)
```

## c = 8

```{r}
#| echo: false
m1s4c3 = rast("../../data/omniscape300/fallrut_RF300_C800_R1000_T50/cum_currmap.tif")
plot(m1s4c3)
```

:::



# Flow potential

The flow potential map shows what connectivity looks like under "null" conditions (i.e., resistance equals 1 for the whole landscape). In other words, flow potential shows what connectivity would look like in the absence of barriers to movement. In the map below, proposed IPCAs are indicated by transparent blue polygons.

## Early winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m2s1c1 = rast("../../data/omniscape300/earlywinter_RF300_C25_R1000_T50/flow_potential.tif")
plot(m2s1c1)
```

## c = 2

```{r}
#| echo: false
m2s1c2 = rast("../../data/omniscape300/earlywinter_RF300_C200_R1000_T50/flow_potential.tif")
plot(m2s1c2)
```

## c = 8

```{r}
#| echo: false
m2s1c3 = rast("../../data/omniscape300/earlywinter_RF300_C800_R1000_T50/flow_potential.tif")
plot(m2s1c3)
```

:::

## Late winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m2s2c1 = rast("../../data/omniscape300/latewinter_RF300_C25_R1000_T50/flow_potential.tif")
plot(m2s2c1)
```

## c = 2

```{r}
#| echo: false
m2s2c2 = rast("../../data/omniscape300/latewinter_RF300_C200_R1000_T50/flow_potential.tif")
plot(m2s2c2)
```

## c = 8

```{r}
#| echo: false
m2s2c2 = rast("../../data/omniscape300/latewinter_RF300_C800_R1000_T50/flow_potential.tif")
plot(m2s2c2)
```

:::

## Summer

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m2s3c1 = rast("../../data/omniscape300/summer_RF300_C25_R1000_T50/flow_potential.tif")
plot(m2s3c1)
```

## c = 2

```{r}
#| echo: false
m2s3c2 = rast("../../data/omniscape300/summer_RF300_C200_R1000_T50/flow_potential.tif")
plot(m2s3c2)
```

## c = 8

```{r}
#| echo: false
m2s3c3 = rast("../../data/omniscape300/summer_RF300_C800_R1000_T50/flow_potential.tif")
plot(m2s3c3)
```

:::


## Fall rut

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m2s4c1 = rast("../../data/omniscape300/fallrut_RF300_C25_R1000_T50/flow_potential.tif")
plot(m2s4c1)
```

## c = 2

```{r}
#| echo: false
m2s4c2 = rast("../../data/omniscape300/fallrut_RF300_C200_R1000_T50/flow_potential.tif")
plot(m2s4c2)
```

## c = 8

```{r}
#| echo: false
m2s4c3 = rast("../../data/omniscape300/fallrut_RF300_C800_R1000_T50/flow_potential.tif")
plot(m2s4c3)
```

:::



# Nomalized current flow

The normalized (cumulative) current flow is calculated as flow potential divided by cumulative current. Values greater than one indicate areas with channelized/bottlenecked flow. Values around 1 (cumulative current ~ flow potential) indicate diffuse current. Values less than 1 indicate impeded flow. In the map below, proposed IPCAs are indicated by transparent blue polygons.

## Early winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m3s1c1 = rast("../../data/omniscape300/earlywinter_RF300_C25_R1000_T50/normalized_cum_currmap.tif")
plot(m3s1c1)
```

## c = 2

```{r}
#| echo: false
m3s1c2 = rast("../../data/omniscape300/earlywinter_RF300_C200_R1000_T50/normalized_cum_currmap.tif")
plot(m3s1c2)
```

## c = 8

```{r}
#| echo: false
m3s1c3 = rast("../../data/omniscape300/earlywinter_RF300_C800_R1000_T50/normalized_cum_currmap.tif")
plot(m3s1c3)
```

:::

## Late winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m3s2c1 = rast("../../data/omniscape300/latewinter_RF300_C25_R1000_T50/normalized_cum_currmap.tif")
plot(m3s2c1)
```

## c = 2

```{r}
#| echo: false
m3s2c2 = rast("../../data/omniscape300/latewinter_RF300_C200_R1000_T50/normalized_cum_currmap.tif")
plot(m3s2c2)
```

## c = 8

```{r}
#| echo: false
m3s2c3 = rast("../../data/omniscape300/latewinter_RF300_C800_R1000_T50/normalized_cum_currmap.tif")
plot(m3s2c3)
```

:::

## Summer

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m3s3c1 = rast("../../data/omniscape300/summer_RF300_C25_R1000_T50/normalized_cum_currmap.tif")
plot(m3s3c1)
```

## c = 2

```{r}
#| echo: false
m3s3c2 = rast("../../data/omniscape300/summer_RF300_C200_R1000_T50/normalized_cum_currmap.tif")
plot(m3s3c2)
```

## c = 8

```{r}
#| echo: false
m3s3c3 = rast("../../data/omniscape300/summer_RF300_C800_R1000_T50/normalized_cum_currmap.tif")
plot(m3s3c3)
```

:::


## Fall rut

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
m3s4c1 = rast("../../data/omniscape300/fallrut_RF300_C25_R1000_T50/normalized_cum_currmap.tif")
plot(m3s4c1)
```

## c = 2

```{r}
#| echo: false
m3s4c2 = rast("../../data/omniscape300/fallrut_RF300_C200_R1000_T50/normalized_cum_currmap.tif")
plot(m3s4c2)
```

## c = 8

```{r}
#| echo: false
m3s4c3 = rast("../../data/omniscape300/fallrut_RF300_C800_R1000_T50/normalized_cum_currmap.tif")
plot(m3s4c3)
```

:::


# Connectivity categories

  1. Impeded (0 - 0.7)
  2. Diffuse (0.7 - 1.3)
  3. Intensified (1.3 - 1.7)
  4. Channelized (1.7 - 9999)

```{r}
m <- c(0, 0.7, 1,
       0.7, 1.3, 2,
       1.4, 1.7, 3,
       1.7, 9999, 4)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
# Early winter
m4s1c1 = as.factor(classify(m3s1c1, rclmat, include.lowest=TRUE))
m4s1c2 = as.factor(classify(m3s1c2, rclmat, include.lowest=TRUE))
m4s1c3 = as.factor(classify(m3s1c3, rclmat, include.lowest=TRUE))
# Late winter
m4s2c1 = as.factor(classify(m3s2c1, rclmat, include.lowest=TRUE))
m4s2c2 = as.factor(classify(m3s2c2, rclmat, include.lowest=TRUE))
m4s2c3 = as.factor(classify(m3s2c3, rclmat, include.lowest=TRUE))
# Summer
m4s3c1 = as.factor(classify(m3s3c1, rclmat, include.lowest=TRUE))
m4s3c2 = as.factor(classify(m3s3c2, rclmat, include.lowest=TRUE))
m4s3c3 = as.factor(classify(m3s3c3, rclmat, include.lowest=TRUE))
# Fall rut
m4s4c1 = as.factor(classify(m3s4c1, rclmat, include.lowest=TRUE))
m4s4c2 = as.factor(classify(m3s4c2, rclmat, include.lowest=TRUE))
m4s4c3 = as.factor(classify(m3s4c3, rclmat, include.lowest=TRUE))
```

## Early winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
plot(m4s1c1, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 2

```{r}
#| echo: false
plot(m4s1c2, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 8

```{r}
#| echo: false
plot(m4s1c3, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

:::

## Late winter

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
plot(m4s2c1, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 2

```{r}
#| echo: false
plot(m4s2c2, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 8

```{r}
#| echo: false
plot(m4s2c3, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

:::

## Summer

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
plot(m4s3c1, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 2

```{r}
#| echo: false
plot(m4s3c2, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 8

```{r}
#| echo: false
plot(m4s3c3, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

:::


## Fall rut

::: {.panel-tabset}

## c = 0.25

```{r}
#| echo: false
plot(m4s4c1, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 2

```{r}
#| echo: false
plot(m4s4c2, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

## c = 8

```{r}
#| echo: false
plot(m4s4c3, col=c('gray95', 'lightskyblue', 'green4', 'red'),
  plg=list(legend=c("1 - Impeded", "2 - Diffuse", "3 - Intensified", "4 - Channelized"), x="bottomright"))
```

:::
