---
title: Covariate data
execute:
  freeze: auto
  echo: true
---

## Introduction

Objectives:

- Prepare raster covariate data

```{r}
library(sf)
library(terra)
library(tidyverse)
```

**Set working directory and folder locations**

```{r}
dropbox <- 'C:/Users/PIVER37/BEACONs Dropbox/pierre vernier/prv_gisdata/north_america/climate/worldclim/'
#bpdata <- 'C:/Users/PIVER37/Documents/github/beaconsproject/geopackage_creator/www/bp_datasets.gpkg'
bpdata <- 'H:/Shared drives/Disturbance Mapping (digitizing)/Data/geopackage_creator/bp_datasets.gpkg'
```

**Read boundary**

```{r}
bnd <- st_read('../../data/yt_caribou.gpkg', 'mcp_buff15k')
bnd_4326 <- st_transform(bnd, 4326)
```

## RASTER TEMPLATE (30-m resolution)

First use QGIS to extract NALC2020 by mask and save in 3578

Then read raster and create boundary raster template

```{r}
nalc <- rast('../../data/raster30/nalc2020.tif')
rast30m <- subst(nalc, 1:19, 1)
plot(nalc)
```

## LANDCOVER

### NALC 2020: forest, wetlands

```{r}
forest <- ifel(nalc %in% c(1,2,5,6), 1, 0) |>
  mask(rast30m)
names(forest)='forest'
conif <- ifel(nalc %in% c(1,2), 1, 0) |>
  mask(rast30m)
names(conif)='conifer'
decid <- ifel(nalc==5, 1, 0)
names(decid)='broadleaf'
mixed <- ifel(nalc==6, 1, 0)
names(mixed)='mixedwood'
wetland <- ifel(nalc==14, 1, 0) |>
  mask(rast30m)
names(wetland)='wetland'
shrub <- ifel(nalc %in% c(8,11), 1, 0) |>
  mask(rast30m)
names(shrub)='shrubland'
grass <- ifel(nalc %in% c(10,12), 1, 0) |>
  mask(rast30m)
names(grass)='grassland'
barren <- ifel(nalc %in% c(13,16), 1, 0) |>
  mask(rast30m)
names(barren)='barren'
snowice <- ifel(nalc==19, 1, 0)
names(snowice)='snowice'
#Combine and save
landcov <- c(forest, conif, decid, mixed, wetland, shrub, grass, barren, snowice)
writeRaster(landcov, '../../data/raster30/landcov.tif', overwrite=TRUE)
plot(landcov)
```

## DISTURBANCES - HUMAN

### Distance to linear disturbances

```{r}
line <- st_read(bpdata, 'sd_line') |>
  st_intersection(bnd) |>
  mutate(id=NULL)
st_write(line, '../../data/bp_covars.gpkg', 'linear_disturbance', delete_layer=TRUE)
plot(st_geometry(line))
dist2line <- rasterize(vect(line), rast30m) |>
  distance() |>
  mask(rast30m)
writeRaster(dist2line, '../../data/raster30/dist2line.tif', overwrite=TRUE)
plot(dist2line)
```

### Distance to areal disturbances

```{r}
poly <- st_read(bpdata, 'sd_poly') |>
  st_intersection(bnd) |>
  mutate(id=NULL)
st_write(poly, '../../data/bp_covars.gpkg', 'areal_disturbance', delete_layer=TRUE)
plot(st_geometry(poly))
dist2poly <- rasterize(vect(poly), rast30m) |> 
  distance() |>
  mask(rast30m)
writeRaster(dist2poly, '../../data/raster30/dist2poly.tif', overwrite=TRUE)
plot(dist2poly)
```

### Distance to linear and areal disturbances

```{r}
linep <- st_buffer(line, 1)
linepoly <- rbind(linep, poly)
st_write(linepoly, '../../data/bp_covars.gpkg', 'all_disturbance', delete_layer=TRUE)
plot(st_geometry(linepoly))
dist2linepoly <- rasterize(vect(linepoly), rast30m) |>
  distance() |>
  mask(rast30m)
writeRaster(dist2linepoly, '../../data/raster30/dist2linepoly.tif', overwrite=TRUE)
plot(dist2linepoly)
```

## NEIGHBOURHOOD VARIABLES

## DISTURBANCES - NATURAL

### Fires

```{r}
# Fires: clip and rasterize
fires <- vect(bpdata, 'fires') |>
  crop(vect(bnd)) |>
  rasterize(rast30m) |>
  mask(rast30m)
fires <- subst(fires, NA, 0) |>
  mask(rast30m)
names(fires) <- "fires"
writeRaster(fires, '../../data/raster30/fires.tif', overwrite=TRUE)
plot(fires)
```

## TOPOGRAPHY

```{r}
# Read, resample, and clip ccdem
dem <- rast('C:/Users/PIVER37/BEACONs Dropbox/pierre vernier/prv_gisdata/canada/dem/can3d30.asc')
dem1 <- crop(dem, bnd_4326)
dem2 <- mask(dem1, bnd_4326)
elev <- project(dem2, nalc, res=30)
names(elev) <- 'elevation'
slope <- terrain(elev, v="slope", neighbors=8, unit="degrees")
names(slope) <- "slope"
rough <- terrain(elev, v="roughness", neighbors=8)
names(rough) <- "roughness"
aspect <- terrain(elev, v="aspect", neighbors=8)
names(aspect) <- "aspect"
north <- cos(aspect * pi / 180)
names(north) <- "northness"
east <- sin(aspect * pi / 180)
names(east) <- "eastness"
topo <- c(elev, slope, aspect, rough, north, east)
writeRaster(topo, '../../data/raster30/topo.tif', overwrite=TRUE)
plot(topo)
```

## BIOCLIMATE (1-km resolution)

```{r}
# Uncomment if downloading for first time
#d1 <- worldclim_tile(var='bio', lon=-128, lat=59, path='../../data/tmp')
#d2 <- worldclim_tile(var='bio', lon=-128, lat=61, path='../../data/tmp')

# Uncomment if previously downloaded
t1 <- rast(paste0(dropbox, 'tile_14_wc2.1_30s_bio.tif'))
names(t1) <- paste0("bio", c(1:19))
t2 <- rast(paste0(dropbox, 'tile_2_wc2.1_30s_bio.tif'))
names(t2) <- paste0("bio", c(1:19))

# Clip to study area
t1_crop <- crop(t1, bnd_4326)
t1_msk <- mask(t1_crop, bnd_4326)
t2_crop <- crop(t2, bnd_4326)
t2_msk <- mask(t2_crop, bnd_4326)
tt <- merge(t1_msk, t2_msk)
tt_3578 <- project(tt, "EPSG:3578")
writeRaster(tt_3578, '../../data/raster1000/bio.tif', overwrite=TRUE)

# Plot bioclimate data
plot(tt_3578)
```
