---
title: Extract covariate values
format: gfm
execute:
  freeze: auto
  echo: true
---

## Introduction

Objectives:

- Extract raster covariate values

```{r}
library(sf)
library(terra)
library(tidyverse)
set.seed(5228)
```

## Presence data

**Prepare presence and availability points**

```{r}
gps <- st_read('../../data/yt_caribou.gpkg', 'gps') |>
  mutate(individual=ID, occurrence=1, ID=NULL) |>
  filter(!season=='other')
glimpse(gps)
```

**Calculate proportion of locations for each individual caribou**

```{r}
gps_id_prop <- group_by(gps, individual) |>
  summarize(n=n()) |>
  mutate(freq=n/sum(n)) |>
  pull()
gps_id_prop
```

**Read predictor rasters and concatenate them**

```{r}
landcov <- rast('../../data/raster30/landcov.tif')
topo <- rast('../../data/raster30/topo.tif')
fires <- rast('../../data/raster30/fires.tif')
dist2line <- rast('../../data/raster30/dist2line.tif')
names(dist2line) <- "dist2line"
dist2poly <- rast('../../data/raster30/dist2poly.tif')
names(dist2poly) <- "dist2poly"
rasters <- c(landcov, topo, fires, dist2line, dist2poly)
names(rasters)
rasters
```

**Extract raster values for GPS locations**

```{r}
pres_vars <- terra::extract(rasters, gps) |>
  mutate(ID=NULL)
pres_vars_sf <- cbind(gps, pres_vars)
pres_vars_sf
glimpse(pres_vars_sf)
```

## Availability data

**Create availability points and extract raster values**

```{r}
npoints <- nrow(pres_vars)
nseasons <- c(rep('earlywinter',length(gps$season[gps$season=='earlywinter'])),
              rep('latewinter',length(gps$season[gps$season=='latewinter'])),
              rep('summer',length(gps$season[gps$season=='summer'])),
              rep('fallrut',length(gps$season[gps$season=='fallrut'])))
avail_vars <- spatSample(rasters, size=npoints, method="random", as.df=TRUE, values=TRUE, xy=TRUE, na.rm=TRUE)
avail_vars_sf <-  st_as_sf(avail_vars, coords = c("x", "y"), crs=3578) |>
  mutate(occurrence=0, x=NULL, y=NULL) |>
  rename(geom=geometry) |>
  mutate(season=nseasons)
avail_vars_sf
glimpse(avail_vars_sf)
```

Add individual IDs using proportion of locations for each individual caribou as weights (prob)

```{r}
avail_ids = sample(unique(gps$individual), nrow(avail_vars_sf), replace=T, prob=gps_id_prop)
avail_vars_sf$individual <- avail_ids
table(avail_ids)
```

## Combine presence and availability data

```{r}
gps_vars <- bind_rows(pres_vars_sf, avail_vars_sf) %>%
  mutate(timestamp=NULL) # id=1:nrow(.)
st_write(gps_vars, '../../data/yt_caribou.gpkg', 'gps_vars', delete_layer=TRUE)
glimpse(gps_vars)
head(gps_vars)
```